Flattering Advice: Avoiding Disappointment in Advice-Giving

### 2025-01-28 manuscript/manuscript.qmd

People often go to great lengths to avoid unpleasant information,
deriving belief-based utility from maintaining favorable views of
themselves [@Golman2017; @Ho2018; @Loewenstein2018]. Moreover, they
often prefer advisers who withhold bad news from them [@Shalvi2019] and
punish those who communicate bad news [@John2019]. Advisers,
anticipating these preferences and consequences face a dilemma: Should
they provide the accurate but potentially painful truth, or shade their
recommendations to preserve the advisee's mood and the adviser's own
standing? Advisers who fear they will be blamed or disliked for
undermining someone's confidence may reasonably avoid candid feedback,
particularly if they don't incur a cost when the advisee makes a
mistake.

In our experiments, however, advisees have no opportunity to punish the
adviser, and advisers' financial incentives are linked to the decisions
the advisees make. We propose that advisers nonetheless have reason to
provide flattering advise because they recognize the psychological toll
of disappointing news. As a result, they may similarly incur a hedonic
cost for delivering unpleasant information that they anticipate may
distress the advisee. Thus, even in the anonymized context of an online
experiment, we hypothesize (and find) that advisers are relucant to
disappoint advisees and advice is thus biased upward, away from accuracy
and toward confirming advisees' priors. We show that this reduces the
quality of advice—but that advisees nonetheless view flattering advisers
more favorably.

>>>>

People go to great lengths to avoid unpleasant information,
People often go to great lengths to avoid unpleasant information,
deriving belief-based utility from maintaining favorable views of
themselves [@Golman2017; @Ho2018; @Loewenstein2018]. Moreover, they
often prefer advisers who withhold bad news from them [@Shalvi2019] and
punish those who communicate bad news [@John2019]. Advisers,
anticipating these preferences and consequences face a dilemma: Should
they provide the accurate but potentially painful truth, or shade their
recommendations to preserve the advisee's mood and the adviser's own
standing? Advisers who fear they will be blamed or disliked for
undermining someone's confidence may reasonably avoid candid feedback,
particularly if they don't incur a cost when the advisee makes a
mistake.

In our experiments, however, advisees have no opportunity to punish the
adviser, and advisers' financial incentives are linked to the decisions
the advisees make. We propose that advisers nonetheless have reason to
provide flattering advise because they recognize the psychological toll
of disappointing news. As a result, they may similarly incur a hedonic
cost for delivering unpleasant information that they anticipate may
distress the advisee. Thus, even in the anonymized context of an online
experiment, we hypothesize (and find) that advisers are relucant to
disappoint advisees and advice is thus biased upward, away from accuracy. 
We show that this reduces the
disappoint advisees and advice is thus biased upward, away from accuracy
and toward confirming advisees' priors. We show that this reduces the
quality of advice—but that advisees nonetheless view flattering advisers
more favorably.
